On exponential sums arising from the classical doubling method

### 2025-03-20 paper.tex
Some cleanup

Hence for a torus character pair $(T,\theta)$ for $G$, we may conjugate $\algebraicGroup{T}$ to a fixed torus $\algebraicGroup{T}'$ and associate to $\theta$ a twisted Frobenius (twisted by an element of $W$ in the twisted Weyl group conjugacy class corresponding to $T$) stable element $x_{T,\theta}$ of $X^*(\algebraicGroup{T}')_{(\mathbb{Q}/\mathbb{Z})_{p'}}$ via the equality
\[
    \exp(\langle x_{T,\theta},y\rangle) = \theta(\aFieldNorm_{\algebraicGroup{T}(\finiteFieldExtension{n}):\algebraicGroup{T}(\finiteField)}(y(\zeta_n)))
\]
for large enough $n$, where $\aFieldNorm_{\algebraicGroup{T}(\finiteFieldExtension{n}):\algebraicGroup{T}(\finiteField)}(t) = t\cdot \Frobenius(t)\cdot \Frobenius^2(t)\cdot  \hdots \cdot \Frobenius^{n-1}(t)$ is the torus norm map. This $x_{T,\theta}$ is well defined up to the Weyl group, and in fact, by \cite[Corollary 13.9]{DigneMichel2020} there is a correspondence between $(T,\theta)$ pairs and $\Frobenius$-stable Weyl group orbits in $X^*(\algebraicGroup{T}')_{(\mathbb{Q}/\mathbb{Z})_{p'}}$.  By composing with \eqref{eq:DL}, we associate to a torus character pair a Frobenius stable semisimple conjugacy class in $G^*$.

\begin{definition}
    Denote the Frobenius semisimple conjugacy class associated to a torus character pair $(T,\theta)$ by $(s_{T,\theta})$. Similarly denote the Frobenius stable Weyl group element of $X^*(\algebraicGroup{T})_{(\mathbb{Q}/\mathbb{Z})_{p'}}$ corresponding to $(T,\theta)$ by $(x_{T,\theta}).$

    The pairs $(T',\theta'),(T,\theta)$ are said to be \emph{geometrically conjugate} is the geometric conjugacy classes $(s_{T,\theta})$ and $(s_{T,\theta'})$ are equal.

    The \emph{Lusztig series} $\LusztigSeries{G}{s}$ associated to a Frobenius stable geometric conjugacy class $(s) \subset \algebraicGroup{G}^*$ consists of the (equivalence classes of the) irreducible representations $\pi$ which appear with non-zero multiplicity in $R_{T, \theta}$ for some $(T,\theta)$ such that $(s_{T,\theta}) = (s)$.
\end{definition}

@@

\begin{definition}
	A \emph{stable function} on $\algebraicGroup{G}(\finiteField)$ is a class function $f$ such that $\sum_{g \in G} f(g)\pi(g)$ is a scalar which only depends on the Lusztig packet for $\pi$. \calvin{FUNCTORIAL TRANSFER FOR REDUCTIVE GROUPS AND CENTRAL
COMPLEXES has this definition}
\end{definition}

A basic example of a class function which doesn't depend of geometric conjugacy 
\begin{proposition}[FIND SOMEWHERE TO CITE or put in appendix]
\label{prop:central_char}
    Any function $f$ supported on the center $\algebraicGroup{Z}(\mathbb{F})$ of $\algebraicGroup{G}(\finiteField)$ is a stable function. 
\end{proposition}
\begin{proof}
\calvin{can probably put this proof in an appendix}
Note it is enough to prove the statement for $f = 1_z$ an indicator function of an element $z \in \algebraicGroup{Z}(\mathbb{F})$. The he proof follows from the two claims:\\
Claim 1: $z$ acts by the scalar $\theta(z)$ in the virtual representation $R_{T,\theta}$.\\
Claim 2: If $(T,\theta),(T',\theta')$ are geometrically conjugate then $\theta|_{\algebraicGroup{Z}(\mathbb{F})} = \theta'_{\algebraicGroup{Z}(\mathbb{F})}$.\\
Proof of claim 1: The representation $R_T^G(\theta)$ is equal to $M \otimes_{\mathbb{C}[T]} \theta$ where $M$ is the virtual module corresponding to the Euler characteristic of cohomology on $X$ which as a variety is a subvariety of $\algebraicGroup{G}$ with the $G \times T$ action via $(g,\ell).x = gx\ell$. The center $\algebraicGroup{Z}$ then commutes across $X$ so the action of $(z,1)$ equals the action of $(1,z)$. The result is on the $\theta$ component of $M$, $z$ acts by $\theta(z)$.\\
Proof of claim 2: Geometric conjugacy is equivalent to for large enough $n$, there exists $g \in \algebraicGroup{G}(\mathbb{F}_n)$ \elad{Here you mean $\algebraicGroup{G}\left(\finiteFieldExtension{n}\right)$?}\calvin{oops yes} such that $g\algebraicGroup{T}g^{-1} = \algebraicGroup{T}'$ and $g(\theta \circ Nm) g^{-1} = (\theta' \circ Nm)$. So when restricted to the center, $\theta \circ Nm = \theta' \circ Nm$. Since $Nm$ is surjective onto the $\mathbb{F}$-points, $\theta|_{\algebraicGroup{Z}(\mathbb{F})} = \theta'|_{\algebraicGroup{Z}(\mathbb{F})}$.
\end{proof}

The main result of this paper is 
\begin{theorem}
\label{thm:Phi_stable}
	$\Phi_{\chi}(g)$ is a stable function. In other words $\dblVirtualJacobiSumScalar{\pi}{\chi}$ only depends on the Lusztig packet of $\pi$.
	%Equivalently	$\frac{\dblVirtualJacobiSumScalar{R_{T, \theta}}{\chi}}{R_{T,\theta}(1)}$ only depends on the geometric conjugacy class of $(T,\theta)$.
\end{theorem}
In the $\GL_n$ case this follows from \Cref{thm:gln-doubling-gauss-sum-in-terms-of-kondo} along with \calvin{Hasse-Davenport? Something else?}.
In this section we prove a slightly weaker result:

\begin{lemma}
\label{lem:indep_geo_conj}
    $\frac{\dblVirtualJacobiSumScalar{R_{T, \theta}}{\chi}}{R_{T,\theta}(1)}$ only depends on the geometric conjugacy class of $(T,\theta)$.
\end{lemma}

@@

\subsection{Computation for Lusztig series}

>>>>

Hence for a torus character pair $(T,\theta)$ for $G$, we may conjugate $\algebraicGroup{T}$ to a fixed torus $\algebraicGroup{T}'$ and associate to $\theta$ a twisted Frobenius (twisted by an element of $W$ in the twisted Weyl group conjugacy class corresponding to $T$) stable element $x_{T,\theta}$ of $X^*(\algebraicGroup{T}')_{(\mathbb{Q}/\mathbb{Z})_{p'}}$ via the equality
\[
    \exp(\langle x_{T,\theta},y\rangle) = \theta(\aFieldNorm_{\algebraicGroup{T}^{\Frobenius^n}:\algebraicGroup{T}^{\Frobenius}}(y(\zeta_n)))
\]
for large enough $n$, where $\aFieldNorm_{\algebraicGroup{T}^{\Frobenius^n}:\algebraicGroup{T}^{\Frobenius}}(t) = t\cdot \Frobenius(t)\cdot \Frobenius^2(t)\cdot  \hdots \cdot \Frobenius^{n-1}(t)$ is the torus norm map. This $x_{T,\theta}$ is well defined up to the Weyl group, and in fact, by \cite[Corollary 13.9]{DigneMichel2020} there is a correspondence between $(T,\theta)$ pairs and $\Frobenius$-stable Weyl group orbits in $X^*(\algebraicGroup{T}')_{(\mathbb{Q}/\mathbb{Z})_{p'}}$.  By composing with \eqref{eq:DL}, we associate to a torus character pair a Frobenius stable semisimple conjugacy class in $G^*$.

    Denote the Frobenius semisimple conjugacy class associated to a torus character pair $(T,\theta)$ by $(s_{T,\theta})$. Similarly denote the Frobenius stable Weyl group element of $X^*(\algebraicGroup{T})_{(\mathbb{Q}/\mathbb{Z})_{p'}}$ corresponding to $(T,\theta)$ by $(x_{T,\theta}).$

We say that the pairs $(T,\theta)$ and $(T',\theta')$ are \emph{geometrically conjugate} if the geometric conjugacy classes $(s_{T,\theta})$ and $(s_{T',\theta'})$ are equal. This is equivalent to the following: there exists $n \ge 1$ and $h \in {\algebraicGroup{G}}^{\Frobenius^n}$ such that $h \algebraicGroup{T}^{\Frobenius^n} h^{-1} = \left(\algebraicGroup{T}'\right)^{\Frobenius^n}$ and such that $$\theta\left(\aFieldNorm_{\algebraicGroup{T}^{\Frobenius^n}:\algebraicGroup{T}^{\Frobenius}}\left(t\right)\right) = \theta'\left(\aFieldNorm_{\left(\algebraicGroup{T}'\right)^{\Frobenius^n}:\left(\algebraicGroup{T}'\right)^{\Frobenius}}\left(h t h^{-1}\right)\right),$$ for any $t \in \algebraicGroup{T}^{\Frobenius^n}$.

The \emph{geometric Lusztig series} $\LusztigSeries{G}{s}$ associated to a Frobenius stable geometric conjugacy class $(s) \subset \algebraicGroup{G}^*$ consists of all (equivalence classes of) irreducible representations $\pi$ which appear with non-zero multiplicity in $R_{T, \theta}$ for some $(T,\theta)$ such that $(s_{T,\theta}) = (s)$.

@@

Let $f \colon \algebraicGroup{G}\left(\finiteField\right) \to \cComplex$ be a class function. Then by Schur's lemma for any irreducible representation $\pi$, there exists $\gamma_{f} \in \cComplex$ such that $$\sum_{g \in G} f(g)\pi(g) = \gamma_{f}\left(\pi\right) \cdot \idmap_{\pi}.$$

\begin{definition}
	We say that  $f \colon \algebraicGroup{G}\left(\finiteField\right) \to \cComplex$ is a \emph{stable function} on $\algebraicGroup{G}(\finiteField)$ if it satisfies the following property: if $\pi$ and $\pi'$ are irreducible representations of $\algebraicGroup{G}(\finiteField)$ such that lie in the same geometric Lusztig series, then $$\gamma_{f}\left(\pi\right) = \gamma_{f}\left(\pi'\right).$$
(See also \cite[Section 4]{LaumonLetellier2023} and \cite[Section 4.1]{ChenBhattacharya2024})
	\end{definition}

Here is a basic example of a stable class function.
\begin{proposition}
\label{prop:central_char}
    If $f \colon \algebraicGroup{G}\left(\finiteField\right) \to \cComplex$ is supported on the center $\algebraicGroup{Z}(\mathbb{F})$ of $\algebraicGroup{G}(\finiteField)$, then $f$ is a stable function.
\end{proposition}
\begin{proof}
By \cite[Lemma 2.2]{Malle2007}, if $\pi$ and $\pi'$ lie in the same geometric Lusztig series, then they have the same central character. The result now immediately follows.
\end{proof}

The main result of this paper is 
\begin{theorem}
\label{thm:Phi_stable}
	The assignment $H \to \cComplex$ given by $$h \mapsto \begin{dcases}
		\chi\left(\detQuadratic\left(\idmap_{\hermitianSpace} - h\right)\right) & h \in G \text{ and } \detQuadratic\left(\idmap_{\hermitianSpace} - h\right) \ne 0,\\
		0 & \text{otherwise.}
	\end{dcases}$$ is a stable function. Equivalently, the assignment $$\pi \mapsto \dblJacobiSumScalar{\pi}{\chi}$$ is constant on geometric Lusztig series.
\end{theorem}
In the $\GL_n$ case this follows from \Cref{thm:gln-doubling-gauss-sum-in-terms-of-kondo} along with \calvin{Hasse-Davenport? Something else?}.
In this section we prove a slightly weaker result:

\begin{lemma}
\label{lem:indep_geo_conj}
    If $(T,\theta)$ and $(T', \theta)$ are geometrically conjugate, then $$\frac{\dblVirtualJacobiSumScalar{R_{T, \theta}}{\chi}}{R_{T,\theta}(1)} = \frac{\dblVirtualJacobiSumScalar{R_{T', \theta'}}{\chi}}{R_{T,\theta}(1)}.$$
\end{lemma}

@@

\subsection{Computation for geometric Lusztig series}
